---
title: 'Lecture 3: Extracting and Parsing Unstructured Data'
author: "Professor L. Jason Anastasopoulos"
date: "9/04/2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Unstructured data and webscraping

- Unstructured data is any data source which is not already in a database.

- Many forms of unstructured data including: XML, HTML and JSON.

- We will discuss each of these in turn and how to extract data from each of these sources.

- Extraction of this kind of data from webpages is also called *webscraping*. 


## JavaScript Object Notation (JSON)

- JSON or JavaScript Object Notation is a file format that uses text and keys to transmit data.

- Invented by programmer Douglas Crockford in the early 2000s.

- It is generally a means for browsers to communicate with servers.

## JSON Data Types

- There are 5 types of JSON data types:
1. Numbers
2. Strings (text)
3. Boolean (True/False)
4. Arrays/Matrices
5. Keys

## JSON Sample

- This is what JSON looks like:
https://alexlitel.github.io/congresstweets/data/2019-08-27.json


## Extracting JSON data with R

- There are a number of packages to do this.

- One of the easier ones to use is *jsonlite*. 

- We will focus on using *jsonlite* to extract JSON from a database of congressional tweets.

## Extracting data from Congressional tweets

- "TweetsofCongress" is a massive database of Congressional tweets collected daily and stored in JSON format.

https://github.com/alexlitel/congresstweets

- Tweets by month and day can be found here: 

https://alexlitel.github.io/congresstweets/

## Let's build a database of tweets from yesterday (08/27)

```{r}
library(jsonlite)

# url with JSON database of Congressional tweets from 08/07

url <- 'https://alexlitel.github.io/congresstweets/data/2019-08-27.json'

# read url and convert to data.frame 
congresstweets <- fromJSON(txt=url)
```

## What does this dataframe look like?
```{r}
head(congresstweets)
```


## Other useful sources of JSON data

- Govinfo API has an incredible wealth of information that you can access in JSON format.

- Sign up with the API here: https://github.com/usgpo/api

## Example of information that you can get using the API

- Bill summaries

```{r}
url = "https://api.govinfo.gov/packages/BILLS-115hr1625enr/summary?api_key=DEMO_KEY"

bill.summaries = 
  fromJSON(txt = url)

head(bill.summaries)
```


## Hyper Text Markup Language (HTML) 

- Hyper Text Markup Language or HTML is the markup language that has served as the basis of all websites.

- Webpages are formatted with a series of tags like the following:
```{r}
# <!DOCTYPE html>
# <html>
#  <head>
#    <title>This is a title</title>
#  </head>
#  <body>
#    <p>Hello world!</p>
#  </body>
#</html>
```

- Data can be extracted from between the tags.

## Extensible Markup Language (XML)

- Similar to HTML, but used primarily for data storage purposes like JSON.

- Does NOT have pre-defined tags like XML (hence the eXtensible part).

- Does have syntacical rules which are detailed here: https://www.w3schools.com/xml/xml_syntax.asp 


## Example of XML document syntax and application

- Syntax
```{r}
# <root>
#  <child>
#    <subchild>.....</subchild>
#  </child>
# </root>
```


## Example of XML document syntax and application

- Application: creation of a note in XML.

```{r}
# <?xml version="1.0" encoding="UTF-8"?>
# <note>
#   <to>Tove</to>
#   <from>Jani</from>
#   <heading>Reminder</heading>
#   <body>Don't forget me this weekend!</body>
# </note>
```




## Parsing HTML and XML in R using Wickham's rvest package

- Let's extract bill data from the govinfo repository: https://www.govinfo.gov/bulkdata

- We will use *rvest* to accomplish this



## First we need to figure out what tags we need to extract

- First go to the page we want to extract data from: https://www.govinfo.gov/bulkdata/BILLSUM/116/hr

- Use https://selectorgadget.com/ to figure out where the tags are that you want.

## Next we will use rvest to get all the nodes that we want

```{r}
library(rvest)

url<- "https://www.govinfo.gov/bulkdata/BILLS/115/1/hr/BILLS-115hr1000ih.xml"

download.file(url, destfile = '/Users/jason/Dropbox/Teaching/Classes-2019-2020-UGA/Text Analysis/Lectures/Week 3 - Extracting Unstructured Data/billsums.xml')

bill_text = xml2::read_html('/Users/jason/Dropbox/Teaching/Classes-2019-2020-UGA/Text Analysis/Lectures/Week 3 - Extracting Unstructured Data/billsums.xml')

bill_text %>%
  html_node("body") %>%
  html_text()
```

## Now turn this into a document

```{r}
billtext = bill_text %>%
  html_node("body") %>%
  html_text()

library(quanteda)

billtext.corpus = corpus(billtext)
billtext.token = tokens(billtext.corpus)

billtext.token
```


```